# pre-class
# Maydelis K. Minaya
# Week 04

Make sure you commit this often with meaningful messages. 

### Question 1:

Using a loop, print the integers from 1 to 50. 

``` {r}
my_integer_loop <- 
 { for (j in 1:50)
  print(j)
 }

```

### Question 2:

A.  Using a loop, add all the integers between 0 and 1000.

``` {r}
my_add_loop <- 
 { for (j in 0:1000)
  print(sum(j+1))
}

```


B. Now, add all the EVEN integers between 0 and 1000 (hint: use seq())

``` {r}
my_addeven_loop <- 
 { for (j in seq(from=0, to=1000, by=2))}
  print(sum(j+1))
  
```

C. Now, repeat A and B WITHOUT using a loop.

```{r}
# A:
# using the seq() to generate the numbers

seq.a <-seq(from=0, to=1000, by=1)
sum(seq.a)
# sum is 500500

# B:
seq.b <- seq(from=0, to=1000, by=2)
sum(seq.b)
# sum is 250500

# the seq function is easier for me than doing the for loop, more intuitive and easier to code 

```

### Question 3:

Here is a dataframe of survey data containing 5 questions :

```{r, eval=FALSE}
survey <- data.frame(
                     "participant" = c(1, 2, 3, 4, 5, 6),
                     "q1" = c(5, 3, 2, 7, 11, 0),
                     "q2" = c(4, 2, 2, 5, -10, 99),
                     "q3" = c(-4, -3, 4, 2, 9, 10),
                     "q4" = c(-30, 5, 2, 23, 4, 2),
                     "q5" = c(88, 4, -20, 2, 4, 2)
                     )

survey
str(survey)
# examine structure of the data frame
```
The response to each question should be an integer between 1 and 5. Obviously, we have some bad values in the dataframe. The goal of this problem is to fix them.

A. Using a loop, create a new dataframe called survey.clean where all the invalid values (those that are not integers between 1 and 5) are set to NA.

``` {r}

survey.clean <- survey






```


B. Now, again using a loop, add a new column to the dataframe called “invalid.answers” that indicates, for each participant, how many bad answers they gave.

``` {r}

invalid.answers <-




```





