---
output:
  word_document: default
  pdf_document: default
  html_document: default
---
# pre-class
# Maydelis K. Minaya 
# Week 03- In-Class Work 
Make sure you commit this often with meaningful messages. 


### Getting Started


We will work with the dataset called [gapminder](https://github.com/jennybc/gapminder), this is a cleaned up version from [Gapminder Data](http://www.gapminder.org/data/). Gapminder contains a lot of great data on all of the nations of the world. We first need to install the gapminder package in R. 

``` {r}
# install.packages("gapminder")

```



``` {r}
library(dplyr)
library(gapminder)
gapminder
head(gapminder) # look at a summary of ghe data before we start to play with it!
```



### Pre-Class Problems

Use **dplyr** functions to address the following questions:

1. How many unique countries are represented per continent?
``` {r}

countries <- unique(gapminder$country)
length(countries)

continents <- unique(gapminder$continent)
length(continents)

(length(countries)/ length(continents))

# There are a total of 142 countries represented in the data
# 5 continents
# 28 unique contries represented per continent, this is an average!

? summarize
gapminder

gapminder %>% 
  group_by(continent) %>% 
  summarize(n=n_distinct(country))

# Results
# A tibble: 5 x 2
  #continent     n
 # <fct>     <int>
# 1 Africa       52
# 2 Americas     25
# 3 Asia         33
# 4 Europe       30
# 5 Oceania       2

```

2. Which European nation had the lowest GDP per capita in 1997?

``` {r}

gapminder %>% 
  filter(continent == "Europe" & year == 1997) %>%
  arrange(gdpPercap) %>% 
    head(n=1) # to make a tible to show just the country with the lowest GDP

# Albania had the lowest GDP in 1997

```

3. According to the data available, what was the average life expectancy across each continent in the 1980s?

``` {r}

gapminder$year # to look at data for the year
summary(gapminder$year)
range(gapminder$year)
subset(gapminder, year == 1982)


gapminder %>%
    filter(year == 1982 | year == 1987) %>%
    group_by(continent) %>%
    summarize(mean_lifeExp = mean(lifeExp))

```

4. What 5 countries have the highest total GDP over all years combined?

``` {r}

gapminder %>%
    mutate(gdp = gdpPercap*pop) %>% # this is to change the GDP back
  # gdpPercap is just the GDP divided by the total pop size
  # mutating in r is to add new variables and preserves existing
  # transmute drops existing variables.
    group_by(country) %>%
    summarise(total.GDP = sum(gdp)) %>%
    arrange(desc(total.GDP)) %>% # to arrange in descdending order
    head(5) # to visualize the top 5 countries

# 1. US
# 2. Japan
# 3. China
# 4. Germany
# 5. United Kingdom

```

5. What countries and years had life expectancies of _at least_ 80 years? _N.b. only output the columns of interest: country, life expectancy and year (in that order)._

``` {r}

gapminder %>%
    filter(lifeExp >= 80) %>%
  select (country, lifeExp, year)

```

6. What 10 countries have the strongest correlation (in either direction) between life expectancy and per capita GDP?

``` {r}

gapminder %>%
    group_by(country) %>% # group all the data by country
    summarise(cor = cor(lifeExp, gdpPercap)) %>% # cor(x,y)
    arrange(desc(abs(cor))) %>% # to arrnge it in descending order by correlation
    head(10) # to see the top 10 countries
  
```

7. Which combinations of continent (besides Asia) and year have the highest average population across all countries? _N.b. your output should include all results sorted by highest average population_. With what you already know, this one may stump you. See [this Q&A](http://stackoverflow.com/q/27207963/654296) for how to `ungroup` before `arrange`ing. This also [behaves differently in more recent versions of dplyr](https://github.com/hadley/dplyr/releases/tag/v0.5.0).

``` {r}
gapminder %>%
  group_by(continent, year) %>%
  filter (continent != "Asia") %>%
  summarise(avg.p <- mean(pop)) %>%
  arrange(avg.p) %>%
  head(5)




```


8. Which three countries have had the most consistent population estimates (i.e. lowest standard deviation) across the years of available data? 

``` {r}
gapminder %>%
    group_by(country) %>% # group all the data by country
    summarise(sd1 = sd(pop)) %>% # sd(pop)
    arrange(sd1)%>% # to arrnge it by ascending order, which is the default in r
    head(3) # to see the top 10 countries

```

9. Subset **gm** to only include observations from 1992 and store the results as **gm1992**. What kind of object is this?

``` {r}

gm1992 <- gapminder %>%
  subset(year ==1992)

gm1992

# this object is a tible (142 x 6)

```


10. Which observations indicate that the population of a country has *decreased* from the previous year **and** the life expectancy has *increased* from the previous year? See [the vignette on window functions](https://cran.r-project.org/web/packages/dplyr/vignettes/window-functions.html).

``` {r}
gapminder %>%
   group_by(country, year) %>%
    select (country, continent, year, lifeExp, pop,gdpPercap )

##

```

