# pre-class
# Maydelis K. Minaya
# Week 06 Week 06


Make sure you commit this often with meaningful messages. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Read the source code for each of the following three functions, puzzle out what they do, and then brainstorm better names.

``` {r}

f1 <- function(string, prefix) {
  substr(string, 1, nchar(prefix)) == prefix
}

f1
# testing out the function
# it works!

f2 <- function(x) {
  if (length(x) <= 1) return(NULL)
  x[-length(x)]
}

f2
# testing out the function
# it works!

f3 <- function(x, y) {
  rep(y, length.out = length(x))
}

f3
# testing out the function
# it works!
# always test the functions to check for errors
```

2. Compare and contrast rnorm() and MASS::mvrnorm(). How could you make them more consistent?

``` {r}

library(MASS)
rnorm(mtcars) 
mvrorms()

# This is the mvrnorm function from the MASS package (Venables and Ripley, 2002), with one small modification to facilitate replication of random samples. The aim is to make sure that, after the seed is reset, the first rows of generated data are identical no matter what value is chosen for n. The one can draw 100 observations, reset the seed, and then draw 110 observations, and the first 100 will match exactly. This is done to prevent unexpected and peculiar patterns that are observed when n is altered with MASS package's mvrnorm.
# R documentation 
# https://www.rdocumentation.org/packages/rockchalk/versions/1.8.129/topics/mvrnorm

# rnorm generates multivariate normal random variates in the space X

# https://www.rdocumentation.org/packages/compositions/versions/1.40-2/topics/rnorm

# To make them more consistent, have an option that inputs cerain variables that the use can pre-define
```





3. Use `lapply()` and an anonymous function to find the coefficient of variation (the standard deviation divided by the mean) for all columns in the mtcars dataset. 

``` {r}
mtcars[,1]
df <-mtcars
df


```











4. Use vapply() to:
    a. Compute the standard deviation of every column in a numeric data frame.
    b. Compute the standard deviation of every numeric column in a mixed data frame. (Hint: you’ll need to use vapply() twice.)
